---
title: "Untitled"
author: "Shih-Chi Chen"
date: "3/13/2022"
output:
  word_document: default
  html_document: default
---

### Simulation study 

Let’s simulate the following linear model:
$y_i=\beta_0+\beta_1x_{i1}+\beta_2x_{i2}+\epsilon_i$
where i=1,2,3,....,100  
$\epsilon_i~N(0,\sigma^2)$,and $\sigma^2=4$

First, create the vector of Y. 
The parameters are set as $\beta=\begin{bmatrix} 10 \\2 \\0.5\end{bmatrix}$
$n=1000, \sigma=2$

```{r}
set.seed(001)
n = 1000; p = 2; sigma = 4;

# Betas
b0 = 10
b1 = 2
b2 = 0.5
beta = c(b0, b1,b2)

# Xs
x1 = runif(n, 0, 10)
x2 = rnorm(n, mean = 4, sd = 10)
xmat = cbind(rep(1,n), x1, x2)

# errors
eps = rnorm(n,0,sigma)

# calculate Y using linear model
y = xmat %*% beta + eps

#bind data
df_lm = as.data.frame(cbind(y,xmat))
colnames(df_lm) = c("y", "intercept", "x1", "x2")
head(df_lm)
```
Estimated the $\beta$ and $\epsilon$

The $\beta_s$ can be calculated through two ways: using formulas and using R function.

```{r}
##sample
# Using formula
beta_form = qr.solve(crossprod(xmat)) %*% crossprod(xmat,y)

compare = cbind(beta_form, beta)
colnames(compare)= c("formula", "actual")
compare
```

#LU simulation
The original sigma^2 is 16 and the realized value is 17.76159
```{r}
#LU Simulation
library(Matrix)
lu_function <- function (X, y) {
  XX <- crossprod(X)
  XX.lu <- expand(lu(XX))
  L <- XX.lu$L
  U <- XX.lu$U
  P <- XX.lu$P
  bs <- forwardsolve(L, crossprod(P, crossprod(X, y)))
  betahat <- backsolve(U, bs)
  
  #sigma^2 hat
  sigma2hat=crossprod(y - (X) %*% betahat) / (dim(X)[1] - dim(X)[2]-1)
  
   #return answer
  list(coefficients = betahat, sigma_sqr = sigma2hat)
}

#compare table
lu_sim=lu_function(xmat,y)
compare = cbind(lu_sim$coefficients, beta)
colnames(compare)= c("formula", "actual")
compare

#sigma^2 hat
lu_sim$sigma_sqr
```

#qr simulation
```{r}
#QR Simulation 

qr_function <- function (X, y) {
  #  Get Q and R matrix for X matrix
  qr_mat=qr(X)
  Q_mat <- qr.Q(qr_mat)
  R_mat <- qr.R(qr_mat)
  
  # Solve Rb = Q'y to get the all beta values
  betahat <- as.vector(backsolve(R_mat, crossprod(Q_mat, y)))
  
  # sigma^2 hat
  sigma2hat=crossprod(y - (X) %*% betahat) / (dim(X)[1] - dim(X)[2]-1)
  
  # return answer
  list(coefficients = betahat, sigma_sqr = sigma2hat)
}


#compare table
qr_sim=qr_function(xmat,y)
compare = cbind(qr_sim$coefficients, beta)
colnames(compare)= c("formula",  "actual")
compare

#sigma^2 hat
qr_sim$sigma_sqr
```

#svd simulation
```{r}
#SVD decomposition Simulation
svd_function <- function (X, y) {
  
  #  Get svd for X matrix
  svd_mat=svd(X)
  svd_sing=svd_mat$d
  U_mat <- svd_mat$u
  V_mat <- svd_mat$v
  
  # Solve x = V*(1/d)*u'*y to get the all beta values
  betahat <- as.vector(V_mat%*%diag(1/(svd_sing))%*%t(U_mat)%*%y)
  
  # sigma^2 hat
  sigma2hat=crossprod(y - (X) %*% betahat) / (dim(X)[1] - dim(X)[2]-1)
  
  # return answer
  list(coefficients = betahat, sigma_sqr = sigma2hat)
}

#compare table
svd_sim=svd_function(xmat,y)
compare = cbind(svd_sim$coefficients, beta)
colnames(compare)= c("formula",  "actual")
compare

#sigma^2 hat
svd_sim$sigma_sqr
```

#eigen simulation
```{r}
#create the adjacent matrix A matrix and force it to Symmetric
set.seed(1)
A_adj=matrix(rbinom(n = 10000, size = 1, prob = 0.5),nrow=100)
A_sim=forceSymmetric(A_adj)

##r+ matrix
#ri=p/ri if ri > 0 ; otherwise ri=0
r_plus=0.85/diag(rowSums(A_sim))
#set the inf number to 0
r_plus[which(!is.finite(r_plus))] = 0 

##Z vector
z=matrix(c(NA),nrow=100)

for(i in 1:100) {
  #zi=(1-p)/n if ri > 0
  if (rowSums(A_sim)[i]> 0){
        z[i,]=(0.15/100)
  #zi=1/n otherwise
    } else {
        z[i,]=(1/100)
    }
}

#I matrix
i_100=as.matrix(c(rep(1, 100)),nrow=100)

#P matrix
P=r_plus%*%A_sim+z%*%t(i_100)

eigen_r=eigen(t(P))

#rank of eigen_r 
order(eigen_r$vectors[,1],decreasing = TRUE)
```

```{r}
##function of Iterative eigen-solver on A matrix (power iteration) Simulation

literation = function(A) {
  
  matrixdim = dim(A)[1]
  
  #create the list to input the q vector
  q = list()
  #initial guess of q^0 
  q[[1]] = matrix(c(rep(1 / matrixdim, matrixdim)), nrow = matrixdim)
  
  #iterative q^k and q^(k-1) by using q^k=Aq^(k-1)/norm(Aq^(k-1))
  for (i in 1:10000) {
    #here we using i+1 instead of i because the first position in q is initial value
    q[[i + 1]] = A %*% q[[i]] / norm(A %*% q[[i]], type = c("2"))
    #length for q^k and q^(k-1)
    length_x = norm(q[[i + 1]] - q[[i]], type = c("2"))
    
    #condition for return value of q^k, lambda, iteration times 
    #when length_x<= tolerance 10^-6
    if (10 ^ -3 >= length_x) {
      eigenvector_q=q[[i+1]]
      lamdba = t(q[[i + 1]]) %*% A %*% q[[i]]
      return(list(times=length(q) - 1,lamdba=lamdba,eigenvector_q=eigenvector_q))
    }
  }
}
Asimlit=literation(t(P))

#rank of Asimlit 
order(Asimlit$eigenvector_q,decreasing = TRUE)
```

```{r}
#compare table for eigen solver (power iteration) the largest lamdba
compare_lam = cbind(Asimlit$lamdba,eigen_r$values[1])
colnames(compare_lam)= c("formula", "actual")
compare_lam

```

```{r}
#compare table for eigen solver (power iteration) the eigenvector
compare_vector = cbind(Asimlit$eigenvector_q,abs(eigen_r$vectors[,1]))
colnames(compare_vector)= c("formula", "actual")
compare_vector

#distance between two vectors
norm(Asimlit$eigenvector_q-abs(eigen_r$vectors[,1]),type="2")
```

```{r}
#jacobi Simulation

jac_literation = function(A, telep = 0.85,  maxiter = 1000,
                     tolx = 1e-3) {
  n<- dim(A_sim)[1]
  ##REFERENCE STA 141C HW-2 Notes:
  A_sparse <- Matrix(A_sim, sparse = TRUE)
  outd_s <- rowSums(A_sparse)
  outdinv_s <- ifelse(outd_s > 0, telep / outd_s, 0)
  z <- ifelse(outd_s > 0, 1 - telep, 1) / n
  # G is the R^+ A matrix
  G <- outdinv_s * A_sparse
  #set the initial vector 
  x = rep(1 / n, n)
  #b from Ax = b:
  bi = x
  #Diagonal of the Matrix:
  aii <- 1 - diag(G) - z
  
  ###JACOBI METHOD:
  ##-------------------
  x = rep(1 / n, n)
  #b from Ax = b:
  bi = x
  #Diagonal of the Matrix:
  aii <- 1 - diag(G) - z
  start.time <- Sys.time()
  for (iter in 1:maxiter) {
    bi <- x
    # (I - P^T)x
    Jacobi_dec <- x - x %*% G - sum(x * z)
    # Jacobi update
    x <- bi - Jacobi_dec / aii
    #stop when the length of two vectors < tolx
    if (sum(abs(x - bi)) < tolx)
      break
    }
    if (iter == maxiter) {
      warning(paste("fail to converge in", maxiter, " iterations!"))
    } else {
      print(paste("converged in", iter, "iterations"))
    }
    #normalize
    pgrank <- abs(x / sum(x))
  # output
  return(pgrank)
  }
#Pagerank values for Jacobi:
jc_pgrank <- jac_literation(A_sim, telep = 0.85,  maxiter = 1000,tolx = 1e-3)

#Runtime Calculation:
timing.jcb <- system.time(jac_literation(A_sim, telep = 0.85,  maxiter = 1000,tolx = 1e-3))

#The top 20 URLS according to the pagerank values:
tp20_index_jc <- order(jc_pgrank,decreasing = TRUE)[1:20]
tp20_index_jc
```

```{r}
#gauss Simulation
gauss_literation = function(A, telep = 0.85,  maxiter = 1000,
                     tolx = 1e-3) {
  n<- dim(A)[1]
  ##REFERENCE STA 141C HW-2 Notes:
  A=as.matrix(A)
  A_sparse <- Matrix(A, sparse = TRUE)
  outd_s <- rowSums(A_sparse)
  outdinv_s <- ifelse(outd_s > 0, telep / outd_s, 0)
  z <- ifelse(outd_s > 0, 1 - telep, 1) / n
  # G is the R^+ A matrix
  G <- outdinv_s * A_sparse
  #set the initial vector 
  x0 = rep(1 / n, n)
  #b from Ax = b:
  bi = x0
  #Diagonal of the Matrix:
  aii <- 1 - diag(G) - z
  k = 1
  while(k < maxiter){
  x = x0
  Gauss_dec <- bi - x0 %*% G - sum(x * z)
  x <- bi - Gauss_dec / aii
  if (any(abs(x - x0) < tolx)){
    gs_pgrank<- x
    break
    }
  k = k +1

  for (iter in 1: maxiter){
    x0 = x
   }
  }
  x0 = x
  #normalize
  gs_pgrank <- abs(x0 / sum(x0))
  #output
  return(gs_pgrank)
}

gauss_liter=gauss_literation(A_sim, telep = 0.85,  maxiter = 1000,tolx = 1e-3)

#Runtime Calculation:
timing.gauss <- system.time(gauss_literation(A_sim, telep = 0.85,  maxiter = 1000,tolx = 1e-3))

#rank top 20
tp20_index_gs <- order(gauss_liter, decreasing = TRUE)[1:20]
tp20_index_gs
```
